### 2022-10-03 Milestone2.Rmd
Deleting excel files not pertinent

Milestone 2

```{r, echo=FALSE}
library(readxl)
library(tidyr)
library(lubridate)

```

```{r}
MPX_data <- read.csv("~/251/files/euro_mpx_cases.csv")
```

•	Description of dataset
o	What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
o	How does the dataset relate to the group problem statement and question?

/newpage

•	Import statement
o	NOTE: Please use datasets available in the PHW251 Project Data github repo Links to an external site. (this is important to make sure everyone is using the same datasets)
o	Use appropriate import function and package based on the type of file
o	Utilize function arguments to control relevant components (i.e. change column types, column names, missing values, etc.)
o	Document the import process

/newpage

•	Identify data types for 5+ data elements/columns/variables
o	Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.
o	Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
o	Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type?

/newpage

•	Provide a basic description of the 5+ data elements
o	Numeric: mean, median, range
o	Character:  unique values/categories
o	Or any other descriptives that will be useful to the analysis