### 2022-11-05 MS3.Rmd
Add files via upload

---
title: "Scenario 2 - MPX in the EU"
author: "Dawson Coblin and Jose Lazo"
date: '2022-11-07'
output: pdf_document
---

# Monkeypox (MPX) Infection Rates in the European Union

This report aims to prepare data to analyze and assess the impact of MPV within  
each country of the EU, and understand how case rates may differ by region and 
other various demographic factors. These findings will aid the implementation
and appropriate response to the epidemic to local health departments.

### Legend of Tasks and Milestones

Milestone  | Description of Section
---------- | ----------------------
Milestone 1|Group Agreement
Milestone 2|Data-set Imported into R; Identify Key Data Elements and Types
Milestone 3|Clean Data-set Create Descriptive Statistics and Data Dictionary

Milestone 3 Sub-sections:  
1.Sub-setting Rows and Columns as Needed- Clean individual Data Frames  
2.Summarizing Data and Creating New Vectors, Organize Vectors for Use  
3.Clean Variables and Combined Data-sets  
4.Descriptive Stats of EU Data and Data Dictionary

\newpage

### 1) Milestone 1: Group Agreement

\newpage
### 2) Milestone 2: Data-set Imported into R; Identify Key Data Elements and Types

```{r packages}
library(ggplot2)
library(tidyverse)
library(readr)
library(lubridate)
library(stringr)
library(readxl)
```
```{r}
euro_mpx_cases <- read_csv("files/euro_mpx_cases.csv")
euro_census_stats <- read_csv("files/euro_census_stats.csv")
euro_pop_denominators <- read_csv("files/euro_pop_denominators.csv")
world_country_regions <- read_csv("files/phw251_projectdata-d6f7578ccda8b47b98f4ab30005bbd76acf496ec/world_country_regions.csv")
```
\newpage

### Milestone 3: 1. Sub-setting Rows and Columns as Needed- Clean individual Data Frames

Data Frame Cleaning

```{r}
#Cleaning the cases df
euro_mpx_cases <- euro_mpx_cases %>%
  select( DateRep, CountryExp, CountryCode  ,ConfCases)

#Creating a vector to use for country code/country name and regions in EU
euro_country_vector <- unique(euro_mpx_cases$CountryCode)

#Cleaning the denominators df
euro_pop_denominators <- euro_pop_denominators %>% 
  filter( TIME_PERIOD == 2022) %>%
  rename( CountryCode = geo , total_pop_2022 = 'OBS_VALUE') %>%
  select( CountryCode, total_pop_2022)

euro_pop_denominators <- euro_pop_denominators[euro_pop_denominators$CountryCode 
                                               %in% euro_country_vector, ]

#Clean census data to include only country code, res_pop. The date for the census 
#dataframe is 2011, so the date is removed and not relevant.

euro_census_stats <- euro_census_stats %>%
  group_by(COUNTRY_CODE, RES_POP) %>%
  summarise(census_pop = sum(pop)) %>%
  rename(CountryCode = COUNTRY_CODE)

#Widen census info to allow each country row to have pop information.
euro_census_stats <- 
  pivot_wider(euro_census_stats, names_from = RES_POP, values_from = census_pop)

euro_census_stats$`0-1000`[euro_census_stats$`0-1000` == 0] <- NA
euro_census_stats$`1000-9999`[euro_census_stats$`1000-9999` == 0] <- NA
euro_census_stats$`10000-99999`[euro_census_stats$`10000-99999` == 0] <- NA
euro_census_stats$`100000-199999`[euro_census_stats$`100000-199999` == 0] <- NA
euro_census_stats$`200000-499999`[euro_census_stats$`200000-499999` == 0] <- NA
euro_census_stats$`500000-999999`[euro_census_stats$`500000-999999` == 0] <- NA
euro_census_stats$GE1000000[euro_census_stats$GE1000000 == 0] <- NA
```

\newpage

### Milestone 3: 2. Summarizing Data and Creating New Vectors, Organize Vectors for Use

```{r}
#We then assign counties with a region code in a new column a list "EU_region,"
#and include the country name and code
sr_northern_e<- c(
  "Denmark" , "Estonia", "Finland", "Iceland", "Ireland", "Latvia", "Lithuania",
  "Norway", "Sweden" , "DK" , "EE", "FI", "IE", "IS", "LT", "LU", "NO", "SE"
)
sr_western_e<- c(
  "Austria", "Belgium", "France", "Germany", "Luxembourg", "Netherlands" , 
  "DE", "AT" , "BE" , "FR", "LV", "NL"
)
sr_eastern_e<- c(
  "Bulgaria", "Czechia", "Hungary", "Poland", "Romania", "Slovakia", "BG", "CZ",
  "HU", "PL" , "RO", "SK"
)
sr_southern_e<- c(
  "Croatia", "Greece", "Italy", "Malta", "Portugal", "Slovenia", "Spain", "EL", 
  "ES", "HR", "IT", "MT", "PT", "SI"
)
sr_western_a<- c( 
  "Cyprus" , "CY"
)

#Create a floor month for the date range to aggregate by month
euro_mpx_cases <- euro_mpx_cases %>%
  mutate(floor_month = floor_date( DateRep , "month"))

#Review the aggregate totals as a value of confirmed cases in Europe and list as
#"cases_grouped_monthly"

euro_mpx_cases <- euro_mpx_cases %>% 
  group_by(floor_month, CountryCode) %>% 
  summarize(cases_grouped_monthly = sum(ConfCases))

#We now add the EU_Region vector based on the country name.
euro_mpx_cases <- euro_mpx_cases %>% 
  mutate(EU_region = case_when(
    CountryCode %in% sr_eastern_e ~ "Eastern Europe",
    CountryCode %in% sr_northern_e ~ "Northern Europe",
    CountryCode %in% sr_southern_e ~ "Southern Europe",
    CountryCode %in% sr_western_e ~ "Western Europe",
    CountryCode %in% sr_western_a ~ "Western Asia"
    ))
```

\newpage

### Milestone 3: 3. Clean Variables and Combined Data-sets

```{r}

#Combined the denominations data for 2022 pop with euro_mpx_cases dataset

euro_mpx_cases <-  inner_join( euro_pop_denominators, euro_mpx_cases, 
                                by= "CountryCode")
#Join census data
euro_mpx_cases <- inner_join(euro_census_stats, euro_mpx_cases,
                              by= "CountryCode")
#Rearrange column order
euro_mpx_cases <- relocate(euro_mpx_cases, floor_month, EU_region, 
                          total_pop_2022 , cases_grouped_monthly, 
                          .before = CountryCode)

euro_mpx_cases<- relocate(euro_mpx_cases, CountryCode, .before = total_pop_2022)

```

\newpage

### Milestone 3: 4. Descriptive Stats of EU Data and Data Dictionary

Descriptive stats:
We find that the census information for 7 of the 29 countries in the EU are not
available. These "0" values were replaced with "NA" to prevent graphing issues.

Region Data:
We see that the regions in EU include various numbers of countries, thus having unequal
populations.

"Eastern Europe" : 6 countries, total region population = 89171711   
"Northern Europe": 9 countries, total region population = 37518701   
"Southern Europe": 7 countries, total region population = 133879004   
"Western Asia"   : 1 countries, total region population = 904705   
"Western Europe" : 6 countries, total region population = 190251495

```{r}

visual <- ggplot(data = euro_mpx_cases, 
                 aes(x = EU_region , y = cases_grouped_monthly))+ 
  theme_minimal(base_size = 15)

visual

```

```{r}
#Upload the Data Dictionary for the central data-frame: "euro_mpx_cases"
Data_Dictionary <- read_excel("~/PHW251_Fall2022/251/MS3_data_dictionary.xlsx")
Data_Dictionary

```