### 2022-11-21 demog_example_mpx.Rmd

---
title: "MPX Example"
output: html_notebook
---

## Ecological analysis using census data

This is an example of creating a metric from the census data relating to percent of population within a country/region residing in small cities

You'll probably need to change the join datasets to match the name and the join by columns that you have set up.  But hopefully this enough to give you an idea.

```{r recode_metric, echo=TRUE, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(magrittr)
library(stringr)
library(tidyr)
library(purrr)
library(lubridate)
library(plotly)

euro_census <- read_csv("https://raw.githubusercontent.com/PHW290/phw251_projectdata/main/euro_census_stats_mod.csv") %>%
  rename_all(. %>% tolower() %>% gsub(" |-|/", "_", .))

mpx_cases <- read_csv("https://raw.githubusercontent.com/PHW290/phw251_projectdata/main/euro_mpx_cases_mod.csv") %>%
  rename_all(. %>% tolower() %>% gsub(" |-|/", "_", .)) %>%
  mutate(month = floor_date(daterep, "month")) %>%
  select(-source, daterep, countryexp, country_code = countrycode, confcases, month)

# collapsing the res_pop into three categories, determining the % of pop in each category for each country
# then categorizing each country into a 3-level metric

census_metric <- euro_census%>%
  mutate(res_size_cat = case_when(res_pop %in% c("0-1000", "1000-9999", "10000-99999") ~ "small",
                                  res_pop %in% c("100000-499999") ~ "medium",
                                  res_pop %in% c("500000-999999", "GE1000000") ~ "large",
                                  TRUE ~ "unknown")
         ) %>%
  group_by(country_code, res_size_cat) %>%
  summarize(group_pop = sum(pop, na.rm = T)) %>%
  mutate(country_pop = sum(group_pop, na.rm = T),
         group_pct = group_pop / country_pop) %>%
  ungroup() %>%
  filter(res_size_cat == "small") %>%
  mutate(pct_small = case_when(group_pct >= 0.80 ~ "Greater than 80% in Small Cities",
                                group_pct >= 0.65 ~ "Between 65% and 80% in Small Cities",
                                group_pct < 0.65 ~ "Less than 65% in Small Cities"))

# merging the demographic metric and population data set with case counts and calculating rates by cat and month

res_size <- mpx_cases %>%
  group_by(country_code, month) %>%
  summarize(cases_by_month = sum(confcases, na.rm = T)) %>%
  left_join(., census_metric, by="country_code") %>%
  group_by(pct_small, month) %>%
  summarize(cases_by_monthcat = sum(cases_by_month, na.rm = T),
            cat_pop = max(country_pop, na.rm = T)) %>%
  mutate(size_cat_rate = (cases_by_monthcat / cat_pop) * 1000000) %>%
  ungroup()

fig <- plot_ly(res_size, x = ~month, y = ~size_cat_rate, type = 'scatter', mode = 'lines', linetype = ~pct_small) %>%
  layout(
    title = 'MPX Case Rates in Europe by Percent of Population Living in Small Cities, May-August 2022',
    xaxis = list(title = '', dtick = "M1", tickformat="%b<br>%Y"),
    yaxis = list(title = 'Case Rates per 1M population'),
    legend=list(orientation = 'h')
  )

fig
```